Bioverse: Origins of Life

### 2024-05-28 src/tex/ms.tex
fix file ending problem
fix most detected problems

% Define document class
%\documentclass[twocolumn,twocolappendix,linenumbers]{aastex631}
\documentclass[modern,linenumbers]{aastex631}
\input{preamble}
\begin{document}

@@

\subsection{Origins of Life Scenarios and their Predictions}
\todo[inline]{Present widely discussed OOL scenarios and their predictions on exoplanet observables; derive testable hypotheses.}
\label{sec:ool_scenarios}
In this section, we present some of the most prominent origins of life scenarios and their observational predictions.
We focus on the necessary environmental conditions for the processes and reactions inherent to each scenario, and aim to identify distinct observables that are accessible via present and near-future remote sensing techniques.

@@

\textbf{Prediction:} Biosignatures occur outside the tidal locking zone and at bulk densities consistent with exposed rock..
\todo[inline]{SR: I'm not a priori sold that tidal locking means that no wet-dry cycles occur. you can still have cycling driven by transient changes in instellation due to flares, for example (e.g., https://iopscience.iop.org/article/10.3847/1538-4357/aadfd1/meta). Similarly, I wonder if 3D effects might not give rise to variability (https://iopscience.iop.org/article/10.3847/PSJ/acc9c4/meta).  I argue that it is more robust to establish a correlation between biosignatures and planets which show evidence of continents/land. I think that Ty Robinson in our department has done some work in this area, his papers might be a good starting point. Other papers which look relevant (but with which I am not familiar, as this is not my area): https://academic.oup.com/mnras/article/511/1/440/6501216, https://academic.oup.com/mnras/article/495/1/1/5733176, https://iopscience.iop.org/article/10.3847/1538-3881/aad775/meta, https://iopscience.iop.org/article/10.3847/1538-3881/aaed3a/meta, https://iopscience.iop.org/article/10.3847/1538-3881/ab2df3/meta}

@@

%\subsection{Origins of Life Hypotheses}
...
\begin{figure*}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/hypotheses.pdf}
        \caption{Population-level hypothesis and null hypothesis on UV irradiance derived from the cyanosulfidic scenario.}
        \label{fig:hypotheses}
    \end{centering}
\end{figure*}
Figure~\ref{fig:hypotheses} shows the hypothesis and null hypothesis derived from the predictions of the cyanosulfidic scenario.
...

@@

%\subsection{H1: Life only originates in hydrothermal vents}
%The hydrothermal vent scenario does not allow oceans deep enough to form an impenetrable layer of high-pressure ice on its floor.
%The resulting allowed parameter space is described by a lower limit on the bulk density. \todo[inline]{+ exclude atmospheric signature for water worlds?}
%...
%
%\subsection{H2: Life only originates in subaerial ponds with wet/dry cycles}
%We parametrize the required exposed land surface in this scenario as a lower limit in bulk density that is higher than for H1.
%Further, the tidal locking timescale of the planet may not be smaller than the age of the system.
%...
%
%\subsection{H3: Life only originates on planets with particular UV irradiance}
\todo[inline]{\citet{Guenther2020} relate U-band energy to bolometric flux.}
\todo[inline]{OPTIONAL: ``We further test the scenario of a linear correlation of past UV flux and biosignature occurrence rate.
This test requires the detection of multiple biosignatures.''
~\\
Test for negative correlations as well?}
Our third theoretical experiment is to test the UV irradiance requirement (Sect.~\ref{sec:predictions:uv}) by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}. \todo[inline]{SUKRIT: got a good reference for this?}  %~\citep{RanjanSasselov2016}.?
Our concrete hypothesis shall be that life only occurs on planets that at some point in their history have received such radiation exceeded a minimum flux $F_\mathrm{NUV, min}$.

@@

\begin{figure}
    \script{inhabited_FGKM.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/inhabited_FGKM.pdf}
        \caption{Host stars of all transiting planets, \glspl{EEC}, and inhabited planets. In a sample of \var{N_nautilus} transiting planets, most \glspl{EEC} and all inhabited planets orbit M~dwarfs.}
        \label{fig:inhabited_FGKM}
    \end{centering}
\end{figure}

@@

\begin{figure*}
    \script{hz_nuv_evo.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/hz_nuv_evo.pdf}
        \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} as a function of host star mass. The scatter plot shows age and host star mass of the transiting planets in the synthetic sample with \gls{EEC} highlighted in green.
        Two of the three planet trajectories shown have extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections), our requirement for abiogenesis.}
        \label{fig:hz_nuv_evo}
    \end{centering}
\end{figure*}

@@

\section{Results}
\label{sec:results}
\subsection{Information content in mass-density space}

\subsection{Information content in tidal locking timescale-density space}

\subsection{Correlation of past UV flux and biosignature occurrence}\label{sec:results_uv}

\subsubsection{Semi-analytical approach}
In Section~\ref{sec:met-semianalytical} we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$ respectively.
Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10).
For $n = \var{semian_Nsamp1}$, strong true evidence for $H_\mathrm{1}$ ($H_\mathrm{null}$) can be expected in $\sim \SI{30}{\percent}$ ($\sim \SI{40}{\percent}$) of all random experiments.
In the majority of cases, the outcome of the survey will be inconclusive.
\begin{figure*}
    \script{semian_true_evidence.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/semian_true_evidence.pdf}
        \caption{Probability to obtain true strong evidence. Left: evidence levels for $H_\mathrm{1}$ and $H_\mathrm{null}$ under sample sizes $n = \var{semian_Nsamp1}$ (solid) and $n = \var{semian_Nsamp2}$ (dashed). The vertical lines denote the thresholds for ``strong'' evidence, $BF_{H_i, H_j}$ > 10, and ``extreme'' evidence, $BF_{H_i, H_j}$ > 100. Right: Probability of true strong evidence for $H_\mathrm{1}$ or $H_\mathrm{null}$, whichever is smaller, as a function of sample size $n$.}
        \label{fig:semian_true_evidence}
    \end{centering}
\end{figure*}
The situation improves with larger samples: for $n = \var{semian_Nsamp2}$, \SI{80}{\percent} of random experiments yield true strong evidence under either $H_\mathrm{1}$ or $H_\mathrm{null}$.

The expected resulting evidence further depends on the a priori unknown abiogenesis rate $f_\mathrm{life}$ and on the \gls{NUV} flux threshold.
\begin{figure*}
    \script{semian_true_evidence.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/semian_evidence-grid.pdf}
        \caption{Probability of obtaining true strong evidence for different abiogenesis rates, \gls{NUV} flux thresholds, and sample sizes. For each of these parameters, higher values increase the probability of yielding strong evidence.}
        \label{fig:semian_evidence-grid}
    \end{centering}
\end{figure*}
Figure~\ref{fig:semian_evidence-grid} illustrates this dependency: For very low values of either parameter, samples drawn under the null or alternative hypotheses are indistinguishable and the Bayesian evidence is always low.
Both higher $f_\mathrm{life}$ and higher \gls{NUV} flux thresholds increase the probability of obtaining strong evidence.
Larger sample sizes enable this at lower values of these parameters.

@@

\begin{figure}
    \script{semian_true_evidence.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/semian_selectivity.pdf}
        \caption{Scaling of the probability of obtaining true strong evidence with sample selectivity. Left: Sampling distribution for different selectivity parameters $s$. Right: Resulting \mbox{P(true strong evidence)} (random, uniform $f_\mathrm{life}$, $F_\mathrm{NUV, min}$). Sampling more extreme values of $F_\mathrm{NUV}$ is more likely to yield strong evidence.}
        \label{fig:semian_selectivity}
    \end{centering}
\end{figure}

\subsubsection{Survey simulations with Bioverse}

@@

A representative recovery of the injected biosignature pattern is shown in Figure~\ref{fig:detections_uv}.
There, we assumed an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$ and a minimum \gls{NUV} flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\SI{}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$.
This leads to a distribution of biosignature detections with detections increasingly occurring above a threshold inferred NUV flux.
\begin{figure}
    \script{detections_uv.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/detections_uv.pdf}
        \caption{Recovered biosignature detections in the \gls{NUV} flux -- biosignature occurrence space. The dashed line denotes the threshold \gls{NUV} flux $F_\mathrm{NUV, min}$.}
        \label{fig:detections_uv}
    \end{centering}
\end{figure}

@@

\begin{figure}
    \script{powergrid.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/powergrid.pdf}
        \caption{Statistical power as a function of threshold \gls{NUV} flux and abiogenesis rate. For a given sample size (here: $N=\var{N_nautilus}$), the achieved statistical power of the survey is enhanced for higher values of $f_\mathrm{life}$ and intermediate values of $F_\mathrm{NUV, min}$.}
        \label{fig:powergrid}
    \end{centering}
\end{figure}

@@

\subsection{Constraining power for the origins of life as a function of biosignature location}
\todo[inline]{How does the location of biosignature detections impact the credibility of OOL scenarios?}
\todo[inline]{biosignature on M~dwarf planet vs. FGK: Impact on UV flux requirement?}
\subsubsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux}
We show in Sect.~\ref{sec:results_uv} that the constraining power for testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis is sensitive on the occurrence of life, the value of this threshold flux, the sample size, and the distribution of sampled past \gls{NUV} fluxes.
In particular the last parameter offers an opportunity to optimize the survey strategy: Although constraints on a planet's UV history have generally large uncertainties~\citep[e.g.,][]{Richey-Yowell2023}, the likely maximum flux of a planet can be inferred and used as a proxy.
Sampling more extreme (low and high) values for the maximum flux increases the probability of obtaining true strong evidence for the hypothesis.
For large sample sizes $\gtrsim 200$, this strategy can push this probability into the \SI{90}{\percent} range.
\todo[inline]{and what about the probability of getting true strong evidence for H0?}

\subsection{Contextual support for potential biosignature detections}
\todo[inline]{discuss how planetary context may impact credibility of tentative biosignature detections (e.g., if there is a good fit with a predicted OOL pattern; or the opposite: the planetary context does not fit well to any OOL scenarios)}
As we have shown in Section~\ref{sec:results_uv}, the interplay of \gls{NUV} evolution and \gls{HZ} occupancy strongly favors late spectral types for abiogenesis via the cyanosulfidic scenario.
This strong predicted correlation between stellar spectral type and the occurrence of life can be used to falsify \gls{ool} scenarios.
The constraint of this context is particularly strong if a candidate biosignature is detected on a planet orbiting an earlier type star, i.e., where it is unexpected in the context of this \gls{ool} scenario.

@@

\subsubsection{Diurnal cycles}
\todo[inline]{We assume that tidal locking implies that there are no wet-dry cycles. Challenge this assumption.}

@@

\bibliography{bib,coauthors}

>>>>

%! suppress = GroupedSubSupScript
%! suppress = Ellipsis
% Define document class
%\documentclass[twocolumn,twocolappendix,linenumbers]{aastex631}
\documentclass[modern,linenumbers]{aastex631}
\input{preamble}
%! suppress = LineBreak
\begin{document}

@@

\subsection{Origins of Life Scenarios and their Predictions}\label{sec:predictions}
\todo[inline]{Present widely discussed OOL scenarios and their predictions on exoplanet observables; derive testable hypotheses.}
\label{sec:ool_scenarios}
In this section, we present some of the most prominent origins of life scenarios and their observational predictions.
We focus on the necessary environmental conditions for the processes and reactions inherent to each scenario, and aim to identify distinct observables that are accessible via present and near-future remote sensing techniques.

@@

\textbf{Prediction:} Biosignatures occur outside the tidal locking zone and at bulk densities consistent with exposed rock.
\todo[inline]{SR: I'm not a priori sold that tidal locking means that no wet-dry cycles occur. you can still have cycling driven by transient changes in instellation due to flares, for example (e.g., https://iopscience.iop.org/article/10.3847/1538-4357/aadfd1/meta). Similarly, I wonder if 3D effects might not give rise to variability (https://iopscience.iop.org/article/10.3847/PSJ/acc9c4/meta).  I argue that it is more robust to establish a correlation between biosignatures and planets which show evidence of continents/land. I think that Ty Robinson in our department has done some work in this area, his papers might be a good starting point. Other papers which look relevant (but with which I am not familiar, as this is not my area): https://academic.oup.com/mnras/article/511/1/440/6501216, https://academic.oup.com/mnras/article/495/1/1/5733176, https://iopscience.iop.org/article/10.3847/1538-3881/aad775/meta, https://iopscience.iop.org/article/10.3847/1538-3881/aaed3a/meta, https://iopscience.iop.org/article/10.3847/1538-3881/ab2df3/meta}

@@

%\subsection{Origins of Life Hypotheses}
...
\begin{figure*}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/hypotheses}
        \caption{Population-level hypothesis and null hypothesis on UV irradiance derived from the cyanosulfidic scenario.}
        \label{fig:hypotheses}
    \end{centering}
\end{figure*}
Figure~\ref{fig:hypotheses} shows the hypothesis and null hypothesis derived from the predictions of the cyanosulfidic scenario.
...

@@

%\subsection{H1: Life only originates in hydrothermal vents}
%The hydrothermal vent scenario does not allow oceans deep enough to form an impenetrable layer of high-pressure ice on its floor.
%The resulting allowed parameter space is described by a lower limit on the bulk density. \todo[inline]{+ exclude atmospheric signature for water worlds?}
%...
%
%\subsection{H2: Life only originates in subaerial ponds with wet/dry cycles}
%We parametrize the required exposed land surface in this scenario as a lower limit in bulk density that is higher than for H1.
%Further, the tidal locking timescale of the planet may not be smaller than the age of the system.
%...
%
%\subsection{H3: Life only originates on planets with particular UV irradiance}
\todo[inline]{\citet{Guenther2020} relate U-band energy to bolometric flux.}
\todo[inline]{OPTIONAL: ``We further test the scenario of a linear correlation of past UV flux and biosignature occurrence rate.
This test requires the detection of multiple biosignatures.''
~\\
Test for negative correlations as well?}
Our third theoretical experiment is to test the UV irradiance requirement (Sect.~\ref{sec:predictions}) by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}. \todo[inline]{SUKRIT: got a good reference for this?}  %~\citep{RanjanSasselov2016}.?
Our concrete hypothesis shall be that life only occurs on planets that at some point in their history have received such radiation exceeded a minimum flux $F_\mathrm{NUV, min}$.

@@

\begin{figure}
    \script{inhabited_FGKM.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/inhabited_FGKM}
        \caption{Host stars of all transiting planets, \glspl{EEC}, and inhabited planets. In a sample of \var{N_nautilus} transiting planets, most \glspl{EEC} and all inhabited planets orbit M~dwarfs.}
        \label{fig:inhabited_FGKM}
    \end{centering}
\end{figure}

@@

\begin{figure*}
    \script{hz_nuv_evo.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/hz_nuv_evo}
        \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} as a function of host star mass. The scatter plot shows age and host star mass of the transiting planets in the synthetic sample with \gls{EEC} highlighted in green.
        Two of the three planet trajectories shown have extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections), our requirement for abiogenesis.}
        \label{fig:hz_nuv_evo}
    \end{centering}
\end{figure*}

@@

\section{Results}
\label{sec:results}
%\subsection{Information content in mass-density space}
%\subsection{Information content in tidal locking timescale-density space}

%\subsection{Correlation of past UV flux and biosignature occurrence}\label{sec:results_uv}

\subsection{Semi-analytical approach}
In Section~\ref{sec:met-semianalytical} we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$ respectively.
Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10).
For $n = \var{semian_Nsamp1}$, strong true evidence for $H_\mathrm{1}$ ($H_\mathrm{null}$) can be expected in $\sim \SI{30}{\percent}$ ($\sim \SI{40}{\percent}$) of all random experiments.
In the majority of cases, the outcome of the survey will be inconclusive.
\begin{figure*}
    \script{semian_true_evidence.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/semian_true_evidence}
        \caption{Probability to obtain true strong evidence. Left: evidence levels for $H_\mathrm{1}$ and $H_\mathrm{null}$ under sample sizes $n = \var{semian_Nsamp1}$ (solid) and $n = \var{semian_Nsamp2}$ (dashed). The vertical lines denote the thresholds for ``strong'' evidence, $BF_{H_i, H_j}$ > 10, and ``extreme'' evidence, $BF_{H_i, H_j}$ > 100. Right: Probability of true strong evidence for $H_\mathrm{1}$ or $H_\mathrm{null}$, whichever is smaller, as a function of sample size $n$.}
        \label{fig:semian_true_evidence}
    \end{centering}
\end{figure*}
The situation improves with larger samples: for $n = \var{semian_Nsamp2}$, \SI{80}{\percent} of random experiments yield true strong evidence under either $H_\mathrm{1}$ or $H_\mathrm{null}$.

The expected resulting evidence further depends on the a priori unknown abiogenesis rate $f_\mathrm{life}$ and on the \gls{NUV} flux threshold.
\begin{figure*}
    \script{semian_true_evidence.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/semian_evidence-grid}
        \caption{Probability of obtaining true strong evidence for different abiogenesis rates, \gls{NUV} flux thresholds, and sample sizes. For each of these parameters, higher values increase the probability of yielding strong evidence.}
        \label{fig:semian_evidence-grid}
    \end{centering}
\end{figure*}
Figure~\ref{fig:semian_evidence-grid} illustrates this dependency: For very low values of either parameter, samples drawn under the null or alternative hypotheses are indistinguishable and the Bayesian evidence is always low.
Both higher $f_\mathrm{life}$ and higher \gls{NUV} flux thresholds increase the probability of obtaining strong evidence.
Larger sample sizes enable this at lower values of these parameters.

@@

\begin{figure}
    \script{semian_true_evidence.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/semian_selectivity}
        \caption{Scaling of the probability of obtaining true strong evidence with sample selectivity. Left: Sampling distribution for different selectivity parameters $s$. Right: Resulting \mbox{P(true strong evidence)} (random, uniform $f_\mathrm{life}$, $F_\mathrm{NUV, min}$). Sampling more extreme values of $F_\mathrm{NUV}$ is more likely to yield strong evidence.}
        \label{fig:semian_selectivity}
    \end{centering}
\end{figure}

\subsection{Survey simulations with Bioverse}

@@

A representative recovery of the injected biosignature pattern is shown in Figure~\ref{fig:detections_uv}.
There, we assumed an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$ and a minimum \gls{NUV} flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\SI{}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$.
This leads to a distribution of biosignature detections with detections increasingly occurring above a threshold inferred NUV flux.
\begin{figure}
    \script{detections_uv.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/detections_uv}
        \caption{Recovered biosignature detections in the \gls{NUV} flux -- biosignature occurrence space. The dashed line denotes the threshold \gls{NUV} flux $F_\mathrm{NUV, min}$.}
        \label{fig:detections_uv}
    \end{centering}
\end{figure}

@@

\begin{figure}
    \script{powergrid.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/powergrid}
        \caption{Statistical power as a function of threshold \gls{NUV} flux and abiogenesis rate. For a given sample size (here: $N=\var{N_nautilus}$), the achieved statistical power of the survey is enhanced for higher values of $f_\mathrm{life}$ and intermediate values of $F_\mathrm{NUV, min}$.}
        \label{fig:powergrid}
    \end{centering}
\end{figure}

@@

\subsection{Constraining power for the origins of life as a function of biosignature location}
\todo[inline]{How does the location of biosignature detections impact the credibility of OOL scenarios?}
\todo[inline]{biosignature on M~dwarf planet vs. FGK: Impact on UV flux requirement?}
\subsubsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux}
We show in Sect.~\ref{sec:results} that the constraining power for testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis is sensitive on the occurrence of life, the value of this threshold flux, the sample size, and the distribution of sampled past \gls{NUV} fluxes.
In particular the last parameter offers an opportunity to optimize the survey strategy: Although constraints on a planet's UV history have generally large uncertainties~\citep[e.g.,][]{Richey-Yowell2023}, the likely maximum flux of a planet can be inferred and used as a proxy.
Sampling more extreme (low and high) values for the maximum flux increases the probability of obtaining true strong evidence for the hypothesis.
For large sample sizes $\gtrsim 200$, this strategy can push this probability into the \SI{90}{\percent} range.
\todo[inline]{and what about the probability of getting true strong evidence for H0?}

\subsection{Contextual support for potential biosignature detections}
\todo[inline]{discuss how planetary context may impact credibility of tentative biosignature detections (e.g., if there is a good fit with a predicted OOL pattern; or the opposite: the planetary context does not fit well to any OOL scenarios)}
As we have shown in Section~\ref{sec:results}, the interplay of \gls{NUV} evolution and \gls{HZ} occupancy strongly favors late spectral types for abiogenesis via the cyanosulfidic scenario.
This strong predicted correlation between stellar spectral type and the occurrence of life can be used to falsify \gls{ool} scenarios.
The constraint of this context is particularly strong if a candidate biosignature is detected on a planet orbiting an earlier type star, i.e., where it is unexpected in the context of this \gls{ool} scenario.

@@

%\subsubsection{Diurnal cycles}
%\todo[inline]{We assume that tidal locking implies that there are no wet-dry cycles. Challenge this assumption.}

@@

%! suppress = MissingBibliographystyle
\bibliography{bib,coauthors}
